SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-10-29 docs/questioning.tex
training draft ready

\usepackage[numbers,sort]{natbib}

@@

\subsubsection{Parse-Tree Validation}

@@

{\sc isr} is a method that contrary to the previous
three, has been applied to counter
different kinds of application attacks~\cite{K09b,KKP03}.
The main idea behind the method, is to
separate code from data by randomizing the legitimate code's 
execution environment. In this way, the malicious code
injected by the attackers who do not know the randomization
algorithm, will not get executed.

@@

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql}
or {\sc xp}ath queries) in various forms, and detect attacks
as outliers from the set of valid code statements.
An early approach, {\it {\sc didafit}}~\cite{LLW02} detects
{\sc sql} injection attacks by recording all database transactions.
Subsequent refinements by Valeur et al.~\cite{VMV05}
tagged each transaction with the corresponding application
as an extension of their anomaly detection framework called
{\it libAnomaly}. {\it {\sc sd}river}~\cite{MS09,MKS09} is
a signature-based mechanism that prevents {\sc sql} and
{\sc xp}ath injection attacks. The signatures generated
during a training phase, are based on features that can
depend either on the code statement or on its execution
environment (i.e. the stack trace).
Then, at runtime, the mechanism checks all statements for compliance
and can block code statements containing injected elements (locus {\sc z}).
% By associating a complete stack trace with the root of an {\sc sql} statement, 
% the mechanism can correlate {\sc sql} statements with their call sites.  
% This increases the specificity of the stored signatures and avoids false alarms.
Furthermore, {\sc amnesia}~\cite{HO05,HO06} is a tool that also detects {\sc sql}
injection attacks by associating a query model with the location of every
{\sc sql} statement within the application. Then at runtime,
monitors the application's execution to detect when {\sc sql} statements
diverge from the expected model.
% dimitro: more mechanisms TBA.

In the case of {\sc xss} attacks the various countermeasures
follow a similar pattern. {\it {\sc swap}}~\cite{WPLKK09}
creates a unique identifier ({\it script {\sc id}}) for
every legitimate script on the server.
Then, a JavaScript detection component placed in a
web proxy (locus {\sc y}) searches for injected scripts
with no corresponding {\sc id} in the server's responses.
If no injected scripts are found, the proxy forwards the
request to the client.
% dimitro: more mechanisms TBA.

@@

    \caption{Comparison summary of mechanisms developed to counter application attacks.}

@@

@@

  \multirow{3}{*}{Code Rewriting}

@@

  \multirow{9}{*}{Training}

>>>>

\usepackage[numbers,sort]{natbib}
\lstset{columns=flexible,showstringspaces=false}

@@

\subsubsection{Parse-Tree Validation}
\label{sec:tree}

@@

{\sc isr} is a method that has been applied to counter
different kinds of application attacks~\cite{K09b,KKP03}.
The main idea behind the it, is to
separate code from data by randomizing the legitimate code's 
execution environment. In this way, the malicious code
injected by the attackers who do not know the randomization
algorithm, will not get executed.

@@

{\it {\sc sm}ask}~\cite{JB07} is a mechanism that detects
both {\sc sql} injection attacks and {\sc xss} attacks.
To do this, it identifies malicious
code by automatically separating user input from legitimate code.
This is done by introducing specific syntactic constructs
that handle server-side languages used for data management
separately.

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql}
or {\sc xp}ath queries) in various forms, and detect attacks
as outliers from the set of valid code statements.
An early approach, {\it {\sc didafit}}~\cite{LLW02} detects
{\sc sql} injection (locus $\Omega$) attacks by recording all database transactions.
Subsequent refinements by Valeur et al.~\cite{VMV05}
tagged each transaction with the corresponding application
as an extension of their anomaly detection framework called
{\it libAnomaly}. {\it {\sc sd}river}~\cite{MS09,MKS09} is
a signature-based mechanism that prevents {\sc sql} and
{\sc xp}ath injection attacks. The signatures generated
during a training phase, are based on features that can
depend either on the code statement or on its execution
environment (i.e. the stack trace).
Then, at runtime, the mechanism checks all statements for compliance
and can block code statements containing injected elements (locus {\sc z}).
Furthermore, {\sc amnesia}~\cite{HO05,HO06,HO05b} is a tool
that also detects {\sc sql} injection attacks (locus {\sc z})
by associating a query model with the location of every
{\sc sql} statement within the application.
Then at runtime, monitors the application's execution to
detect when {\sc sql} statements diverge from the expected model.

In the case of {\sc xss} attacks the various countermeasures
follow a similar pattern. {\it {\sc swap}}~\cite{WPLKK09}
creates a unique identifier ({\it script {\sc id}}) for
every legitimate script on the server.
Then, a JavaScript detection component placed in a
web proxy (locus {\sc y}) searches for injected scripts
with no corresponding {\sc id} in the server's responses.
If no injected scripts are found, the proxy forwards the
request to the client. This mechanism is relatively
inflexible since it does not support dynamic scripts.
In addition it imposes a significant overhead (see Table~\ref{tab:comp}).
The authors of {\it {\sc xssds}}~\cite{JEP08}
have implemented a similar mechanism that also supports
dynamic and external scripts. Specifically,
during training phase they build a list of all benign scripts.
In the case of external scripts they keep a white list
of all the valid domain names that contain scripts
used by the application.

The training subcategory, includes some mechanisms
that can be easily circumvented. For example,
consider an application that will send the password 
for a forgetful user via email by executing
the following query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
In a similar manner, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a mimicry attack like the
one mentioned in Subsubsection~\ref{sec:tree}.

In general, the detection accuracy of training approaches is
heavily based on the coverage that is achieved during the
training phase. If the coverage is insufficient the
production of false alarms is very likely.
In addition, when a code statement is altered, a new
training phase is necessary.

All the above mechanisms are relatively easy to deploy.
{\sc dsl} injection attack countermeasures
can be retrofitted to a system typically by changing
some configuration files. This does not apply
to {\sc amnesia} though, since significant source code
modifications are required for every query that exists
in the application. Finally, {\sc swap} and {\sc xssds}
are implemented within a proxy on the server-side.

@@

    \caption{Comparison summary of mechanisms developed to counter application attacks based on executable source code injection.}

@@

  \multirow{3}{*}{Parse-Tree Validation}

@@

  \multirow{4}{*}{Code Rewriting}

@@

  \multirow{7}{*}{Training}
